\vspace{-2cm}A generative model for urban firm networks

### 2020-11-18 Response_to_reviewers.tex

%% start of file `template.tex'.
%% Copyright 2006-2013 Xavier Danaux (xdanaux@gmail.com).
%
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License version 1.3c,
% available at http://www.latex-project.org/lppl/.

\documentclass[10pt,a4paper,sans]{moderncv}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

\usepackage[document]{ragged2e}
% pour justifier

% moderncv themes
\moderncvstyle{banking}                            % style options are 'casual' (default), 'classic', 'oldstyle' and 'banking'
\moderncvcolor{red}                                % color options 'blue' (default), 'orange', 'green', 'red', 'purple', 'grey' and 'black'
\renewcommand{\familydefault}{\rmdefault} 
%\nopagenumbers{} 
\usepackage[utf8]{inputenc} 
\usepackage[scale=0.9]{geometry}

\def \draft {1}
\usepackage{xparse}
\usepackage{ifthen}
\DeclareDocumentCommand{\comment}{o m o o o o}
{\ifthenelse{\draft=1}{
  \IfValueT{#1}{
      \textcolor{red}{\textbf{C (#1) : }#2}
      \IfValueT{#3}{\textcolor{blue}{\textbf{A1 : }#3}}
      \IfValueT{#4}{\textcolor{green}{\textbf{A2 : }#4}}
      \IfValueT{#5}{\textcolor{red!50!blue}{\textbf{A3 : }#5}}
      \IfValueT{#6}{\textcolor{blue}{\textbf{A4 : }#6}}
    }
    \IfNoValueT{#1}{
      \textcolor{red}{\textbf{C : }#2}
      \IfValueT{#3}{\textcolor{blue}{\textbf{A1 : }#3}}
      \IfValueT{#4}{\textcolor{green}{\textbf{A2 : }#4}}
      \IfValueT{#5}{\textcolor{red!50!blue}{\textbf{A3 : }#5}}
      \IfValueT{#6}{\textcolor{blue}{\textbf{A4 : }#6}}
    }
 }{}
}

\firstname{}
\lastname{}
\begin{document}

% recipient data
\recipient{Editor PONE}{}
\date{\today}
\opening{Dear Editor,}
\closing{Yours faithfully,\\
Juste Raimbault and co-authors\\
University College London
}
         % use an optional argument to use a string other than "Enclosure", or redefine \enclname

%\makelettertitle

\justify

\textbf{Reply to the editor / Journal requirements}

\textit{}

\medskip

$\rightarrow$

\bigskip

\textbf{Reply to reviewers}

\medskip

\textbf{First referee:}

\medskip

%### Summary of the research and your overall impression

%The authors develop a model aiming at capturing the determinants of inter-city firms ownership links formation. The model can be seen as an interesting combination of gravity models and preferential attachment that allows to account for the effect of path-dependence (or more accurately, the effect of circular causation/feedback).
%The theoretical implications of the model are rigorously explored by uncovering the link between the model parameters and some of the city network structural features.
%The model is eventually fitted to empirical data to evaluate the impact of the model variables on the formation of firms ownership links. The results are consistent with the literature where the gravity model has been applied in similar contexts (mainly trade and migration), although this is not explicitly mentioned by the authors. The effect of path dependence is found to be strongly significant.
%The authors then use the calibrated model to evaluate the effect of movement restrictions and change in institutional proximity on the city network structure. While the observed effects seem qualitatively relevant, it is difficult to assess the practical usefulness of this exercise. Because: 1) the authors don't discuss the ability of their model to reproduce the structural features of the empirical network that it was calibrated on, and, 2) the global predictive performance of the model (through the Mean squared error on log and the log of mean squared error) seems insufficient for such an exercise (roughly almost an order of magnitude of difference) seems too large at a first glance.
%In conclusion, the idea of the model is interesting and novel. The developed model appears to be a useful tool to analyze the effects of the chosen variables in the case of inter-city firm ownership links formation. Using it as a predictive tool seems however less convincing.

%### Major issues

\begin{enumerate}
%1.1 -

	\item \textit{The work is at the intersection of three strands of literature: a) spatial interaction models, b) network analysis, and c) preferential attachment. Yet this does not appear in the literature review. The work could gain more depth from a better discussion of the literature it is grounded in.}

	$\rightarrow$

	\medskip

%1.2 -

	\item \textit{The empirical network is sparse (density ~ 0.03). Do the authors consider the absent links in the regression and/or the generative model they propose ? (from the fact that the log is taken in the statistical model and the computation of the mean square error on logs, it appears that it is not the case). This could be made clear and discussed, especially that the absence of links is the norm rather than the exception, meaning that the model only treats only a minor proportion of cases.}

	$\rightarrow$

	\medskip

%1.3 -

	\item \textit{Given that the model is generative, the temporal aspect of the model can be an issue. Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectoral proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model. Can the authors mention (and may be discuss) this issue?}

	$\rightarrow$

	\medskip

%1.4 -

	\item \textit{Application: Two potentially problematic aspects: 1) the MSE on logs and the log MSE appear to be quite large to perform an accurate prediction (for an MSE on logs $\sim 5$, the average multiplicative error can roughly be expected to be around $\exp(5^0.5) \sim 9$, which is almost an order of magnitude of difference; the same quick calculation can be applied to the log MSE). Can the authors discuss this point? (May be using a mean absolute error instead of mean squared error is a more tangible estimate of the error?)}

	$\rightarrow$

	\medskip

	\item \textit{2) The authors do not show how well the model reproduces the metropolization and the communities of the empirical network. If the model does not perform well at reproducing these properties of the network, it can mean that the predictors they used are not relevant to such features. This would imply that the model is of limited relevancy to study the impact of crises on these structural features.}

	$\rightarrow$

	\medskip

%### Minor issues
%2.1 -

	\item \textit{Do you consider self-connections in the network ? These have been shown to represent an important proportion of firm ownership links.}

	$\rightarrow$

	\medskip

% 2.2 - 
	\item \textit{Could you please better justify the choice of indicators (Internationalizationand Metropolization).}

	$\rightarrow$

	\medskip

%2.3 -

	\item \textit{The model is ambiguously described. The link probability update $p_{ij}$ is between 0 and 1, does it mean that 1) at each time step each link ij has a $p_{ij}$ probability to be updated or that 2) like in the preferential attachment model, only one link is updated at each time step (and thus the real probability is $p_ij/sum_{lk}(p_{lk})$)?}

	$\rightarrow$

	\medskip

%2.4

	\item \textit{Statistical analysis: A negative binomial or Gamma regression can also be tried. In the case the absence of links is taken into account in regression, a zero-inflated or hurdle regression can be done.}

	$\rightarrow$

	\medskip

%### Misc:

	\item \textit{For a better clarity, please precise that the model generates a directed network.}

	$\rightarrow$

	\medskip

	\item \textit{Link creation probability formula - industrial proximity: oversight to remove k (should not depend on k).}

	$\rightarrow$

	\medskip

	\item \textit{"Model 4 also indicates the absence of overfitting": Can the authors be more precise?}

	$\rightarrow$

	\medskip

	\item \textit{Model exploration: The variation of community size was explored but the indicator has not been defined.}

\end{enumerate}

\bigskip
\bigskip

\textbf{Second referee:}

\medskip

%This article presents a very interesting and novel approach to identifying corporate city networks, and for testing their effects in terms of economic resilience in cities. The approach is well-supported by previous literature and techniques, and it is clear that the authors have a solid background in statistical analysis.
%My overall assessment is that the paper is high-quality and along the lines of what should be published in PLOS One. Thus rather than focusing on what the paper's merits are, I will point to a number of areas for improvement. If I am not mistaken, PLOS One does not have a word count limit, meaning that -- unlike many other journals -- there are effectively no limitations on the amount of detail the authors can provide.
%My main comment is that this paper would be almost unintelligible to the 'average' reader, even an economic geographer. It is full of jargon (not in a bad way) that is clearly of more use to a network scientist, statistician, or even a physicist with an understanding of basic geography. However, as an economic geographer myself, I must say that I would have liked to see a much greater focus on the actual geography, and theory, behind the analysis. To rectify this, I suggest the following:

\begin{enumerate}

%1.

	\item \textit{Clarify the paper's aims. The part about border closures is interesting but seems almost tacked-on. Is the paper about 'what if' borders closed (e.g. COVID, Brexit), or is it about the communities that exist and the relative hierarchies within the network. To be honest, I don't know if I could pick one after reading it. The conclusion claims that "this paper [is] aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset". This should be reoriented toward a more theoretical framing, and speak to the network geographic ties behind it.}

	$\rightarrow$

	\medskip

%2.
	\item \textit{Clarify the jargon. What is a 'generative model'? What is 'synthetic' data?}

	$\rightarrow$

	\medskip

%3.
	\item \textit{Clarfiy the methods and data. Explain what Amadeus is, how it is gathered etc. Explain the tools that are used to analyse it. Why those statistical packages and not others? It would also be useful to describe the data. How many cities and how many firms?}

	$\rightarrow$

	\medskip

%4.
	\item \textit{And most importantly, clarify the theoretical framing. There is plenty of literature cited, but which of these provided inspiration for this study? For example, is this speaking to the 'global city networks' literature (e.g. Neal, Derudder) or more the economic connectivity literature (e.g. Acemoglu)?}

	$\rightarrow$

	\medskip

%Once these issues are addressed I feel the paper will be in much better shape. There also a few awkward phrases (e.g. 'crow fly' distance) throughout, so I recommend a carefully re-reading prior to resubmission.

\end{enumerate}

\bigskip

\textbf{Third referee:}

\medskip

%It is with great interest that I read this paper, as I feel it has much to contribute to current debates on cities - in particular how to measure specific processes. The authors have done well to think about how the complexity of city processes might be measured, but I feel there are issues that need to be addressed before it is ready to be published in this journal.

\begin{enumerate}
%1)

	\item \textit{The main literature discussion appears to be in the introduction - but this is relatively high level. The research gap that the authors appear to be filling is the operationalisation of the processes (internationalisation, metropolisation, regionalisation and specialisation) within a city. However, it is unclear how this links to the processes you idetnfy as driving the network (listed 1-4 on pg.2) and then the model itself. The processes mentioned appear to be various ways of seeing proximity (ie agglomeration is spatial proximity, and previous linkages could be related to social or institutional proximity) - and whilst link back to previous work mentioned, there needs to be more discussion of past research.}

	$\rightarrow$

	\medskip

%2)
	\item \textit{The processes are mentioned in the methods section - but again exactly what is being measured and what each means is not well related to the process. For example, why exactly are geographic structures captured by internationalisation? Why have you chosen metropolisation to be represented in this way?}

	$\rightarrow$

	\medskip

%3)
	\item \textit{More information on why the synthetic and real world models have been set up are needed. What do they bring to our understandings? You mention how internationalisation, metropolisation, fit within the model .... but what about regionalisation and specialisation? Are these latter just an outcome of the analysis, whilst the former are model parameters? Then why are they treated as the same in the introduction?}

	$\rightarrow$

	\medskip

%4)
	\item \textit{Why in the statistical analysis do you then look at the 'most determining geo-economic factors influencing the creation of links between cities' - there was not real literature review to identify these - how do you then identify them? What is the relationship to the research question? Also pg 6 again describe the processes driving the networks (from pg 2 as noted above) - but these measures where not really discussed in when the model was being explained. How do they fit?}

	$\rightarrow$

	\medskip

%5)
	\item \textit{The application of economic shocks appears to be an additional part of the model that is not properly discussed again in the literature review. I wonder if this is actually a different paper? I felt there were lots of interesting aspects to the research - and thinking about how to divide out the components to allow better discussion for each would benefit the research in that you could dive alot deeper into specific discussions.}

	$\rightarrow$

	\medskip

%I hope these suggestions help the authors develop this work, I feel it is very interesting and worthwhile. Deeper discussion on the different elements would really help to add to the contribution they are making to the literature. Given word limits - this is likely done by rethinking the research into several papers. Good luck with your revisions.

\end{enumerate}

\end{document}

%% end of file `template.tex'.

### 2021-02-23 main.tex
Update on Overleaf.

@@

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, straightening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}, by developing strategic industries for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

@@

In this paper we choose to focus on urban networks of firms to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities within a generative model. The specific question we tackle is the complementarity of different processes to generate such urban networks. The originality of the present paper is to examine these issues within the system of cities framework \cite{berry1964cities}. In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2019arXiv191014652Z}.

@@

The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study. We then proceed to a sensitivity analysis and exploration of the model on synthetic data. The model is calibrated on real European firm ownership data. We finally discuss implications of our results and possible extensions.

@@

The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalized to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described below); (iii) city economic size, which presents agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages (influence of already existing links in the past). This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.

@@

We consider two main macroscopic indicators to quantify the generated network. Firstly, geographical structures are captured by internationalisation (modularity of countries in the network). The situation of a minimised modularity corresponds to dense connections between cities of different countries in Europe and sparse connections between cities of the same European country. In practice, modularity is computed as follows
\begin{equation}
    I = \frac{1}{2 \left|E\right|} \cdot \sum_{i,j} \left(w_{ij} - \frac{d_i d_j}{2 \left|E\right|}\right) \mathbbm{1}_{c_i = c_j}
\end{equation}
where $d_i$ is the weighted degree of city $i$ in country $c_i$.

Secondly, we consider the metropolisation indicator as the correlation between the weighted degree and the GDP of the city. With this indicator we capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities. It is computed with a Pearson correlation estimator as
\begin{equation}
    M = \hat{\rho} \left[d_i , E_i \right]
\end{equation}

@@

@@

%%%%%%%%%%%%%
\begin{figure}
    %\begin{center}
    %    \includegraphics[width=\linewidth]{figures/Fig1.png}
    %\end{center}
    \vspace{2cm}
    \caption{Distributions of network properties. \textit{(Left)} Cumulated distribution for edge weight, with power law fit (red) and log-normal fit (green); \textit{(Right)} Cumulated distribution for weighted degree.}
    \label{fig:nwdist}
\end{figure}
%%%%%%%%%%%%%

@@

The communities reflect characteristic socio-cultural structures and cross-border economic ties throughout Europe. Cities in France, Belgium and Luxembourg belong to the same cluster as they share French as a common business language, but also cross-border tax arrangements creating easier economic interactions \cite{DecovilleDurand2019}. The same socio-cultural proximity effect is seen in cities of the United Kingdom and Ireland, and those of Sweden, Finland, Estonia, Latvia and Lithuania. The cluster composed of Dutch and Polish cities is justified by the fact that firms from the Netherlands are among the first foreign owners of polish firms since 2010 \cite{2019arXiv191014652Z}. The cluster including cities in Germany, Czech Republic, Austria, Slovakia, Hungary and partly in Romania, find its origin in the dependence of the Central and Eastern European economies on German and Austrian large companies mostly from the automotive sector (i.e. Audi, Volkswagen). In addition, all cities of the later cluster have in common to be at the intersection of a very important freight transport route in Europe from Germany to Turkey \cite{Zdanowska2017}. 
%%%%%%%%%%%%%%
\begin{figure}
    %\begin{center}
    %\includegraphics[width=\linewidth]{figures/Fig2.png}
    %\end{center}
    \vspace{2cm}
    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the turnover within the Functional Urban Area, and color the community it belongs to. Urban Areas in communities of size smaller than 3 were not colored for readability.}
    \label{fig:fig2}
\end{figure}
%%%%%%%%%%%%%%

@@

This multiplicative form linearly transformed through logarithms is common practice for spatial interaction models. Note that no constraints are considered, since links include information about the turnover and a share of ownership at the same time. 
Estimation results are given in Table~\ref{tab:reg}. We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}. The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero. Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17. Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanatory power (model 3). The best model among the linear ones is the model 4, with all factors and fixed effects, both in terms of explained variance ($R^2 = 0.31$) and Akaike information criterion. Model 4 also indicates the absence of overfitting. Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations). Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e. similarity is more important than complementarity for this dimension; (iv) a large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.

@@

%%%%%%%%%%%%%
\begin{figure}
    %\begin{center}
    %    \includegraphics[width=\linewidth]{figures/Fig3.png}
    %\end{center}
    \vspace{2cm}
    \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row), with a high (resp. low) gravity decay parameter for the left column (resp. right).\label{fig:fig3}}
\end{figure}
%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%
\begin{figure}
	%\begin{center}
	%    \includegraphics[width=\linewidth]{figures/Fig4.png}
	%\end{center}
	\vspace{2cm}
    \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size. Both plots show a transition from a local to a global regime. (Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important. \label{fig:fig4}}
\end{figure}
%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}
    %\includegraphics[width=\linewidth]{figures/Fig5.png}
    \vspace{2cm}
    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_O$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}
    %\includegraphics[width=\textwidth]{figures/Fig6.png}
    \vspace{2cm}
	\caption{Maximal integration in terms of community size is achieved at an intermediate value of d$_o$: emergence of a regional regime; maximal size depends on the role of sectors $\gamma_S$ in a decreasing way when the origin size is deactivated, and in an increasing way when $\gamma_o=1$; this regime disappears when the influence of the origin size becomes too large.\label{fig:fig6}}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}
	%\begin{center}
    %\includegraphics[width=0.75\linewidth]{figures/Fig8.png}
    %\end{center}
    \vspace{2cm}
	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives. $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models). Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
\end{figure}
%%%%%%%%%%%%%%

>>>>

% Keywords
% Urban firm networks; Simulation model; Model calibration

@@

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, strengthening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}. This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

@@

In this paper we choose to focus on interurban interactions between firms. Firm networks have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020} \cite{BidaRozenblat2020}. Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009} \cite{OCleary2016}. Following this literature the present paper positions itself within the global networks theoretical framework and contributes to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities in terms of firm networks within a generative model. The specific question we tackle is the complementarity of different processes to generate such urban networks. The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017} . In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.

@@

The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study. We then proceed to a sensitivity analysis and exploration of the model on synthetic data. The model is calibrated on empirical European firm ownership data. We finally discuss implications of our results and possible extensions.

@@

The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described below); (iii) city economic size, which presents agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages (influence of already existing links in the past). This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.

@@

We consider two main geographical phenomena and therefore two macroscopic indicators to quantify the generated network. Firstly, geographical structures are captured by internationalisation of city-interactions in the context of intensification of globalisation and transnational city networking at the worldwide level \cite{taylor2001specification}. Hence we define the internationalisation indicator as the modularity of countries in the network. The situation of a minimised modularity corresponds to dense connections between cities of different countries in Europe and sparse connections between cities of the same European country. In practice, modularity is computed as follows
\begin{equation}
    I = \frac{1}{2 \left|E\right|} \cdot \sum_{i,j} \left(w_{ij} - \frac{d_i d_j}{2 \left|E\right|}\right) \mathbbm{1}_{c_i = c_j}
\end{equation}
where $d_i$ is the weighted degree of city $i$ in country $c_i$.

Secondly, we consider the geographical process of metropolisation. The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{sassen1991global}. Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most qualified and dynamic activities \cite{RozenblatPumain2007}. Metropolisation gradients have been already explored in literature in city-systems considering several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}. We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.

The later is computed with a Pearson correlation estimator as
\begin{equation}
    M = \hat{\rho} \left[d_i , E_i \right]
\end{equation}

@@

We base our study on empirical data from \emph {ADAMDEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe. AMADEUS indicates the exact geolocation, turnover and activity codes for owners and owned companies in all sectors in Europe. The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst & Young (EY ), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard & Poor's, Thomson Reuters or Duns and Bradstreet. A representativity study shows that \emph {ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of reality in terms of amount of foreign direct investments in 2011 in Poland \cite{Zdanowska2018}.

@@

%%%%%%%%%%%%%
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{figures/Fig1.png}
    \end{center}
    \vspace{2cm}
    \caption{Distributions of network properties. \textit{(Left)} Cumulated distribution for edge weight, with power law fit (red) and log-normal fit (green); \textit{(Right)} Cumulated distribution for weighted degree.}
    \label{fig:nwdist}
\end{figure}
%%%%%%%%%%%%%

@@

The communities reflect characteristic socio-cultural structures and cross-border economic ties throughout Europe. Cities in France, Belgium and Luxembourg belong to the same cluster as they share French as a common business language, but also cross-border tax arrangements creating easier economic interactions \cite{DecovilleDurand2019}. The same socio-cultural proximity effect is seen in cities of the United Kingdom and Ireland, and those of Sweden, Finland, Estonia, Latvia and Lithuania. The cluster composed of Dutch and Polish cities is justified by the fact that firms from the Netherlands are among the first foreign owners of polish firms since 2010 \cite{2020Zdanowska}. The cluster including cities in Germany, Czech Republic, Austria, Slovakia, Hungary and partly in Romania, find its origin in the dependence of the Central and Eastern European economies on German and Austrian large companies mostly from the automotive sector (i.e. Audi, Volkswagen). In addition, all cities of the later cluster have in common to be at the intersection of a very important freight transport route in Europe from Germany to Turkey \cite{Zdanowska2017}. 
%%%%%%%%%%%%%%
\begin{figure}
    \begin{center}
    \includegraphics[width=\linewidth]{figures/Fig2.png}
    \end{center}
    \vspace{2cm}
    \caption{{\bf Map of network communities obtained by modularity maximisation.} The area of circles gives the turnover within the Functional Urban Area, and color the community it belongs to. Urban Areas in communities of size smaller than 3 were not colored for readability.}
    \label{fig:fig2}
    % TODO redo map with open resources (see editor request) https://www.naturalearthdata.com/downloads/110m-cultural-vectors/
\end{figure}
%%%%%%%%%%%%%%

@@

This multiplicative form linearly transformed through logarithms is common practice for spatial interaction models. Note that no constraints are considered, since links include information about the turnover and a share of ownership at the same time. Estimation results are given in Table~\ref{tab:reg}. We also fit a Poisson generalized linear model with a logarithmic link on the rounded weights \cite{flowerdew1988fitting}. The findings show that when distance only is taken into account (model 1), it has a significant negative effect on link creation, but with an explained variance close to zero. Model 2 adds fixed effects by pair of countries, for which around a third are significant, and with a better $R^2$ of 0.17. Adding the economic size at origin and destination (turnovers $T_i$ and $T$, model 3) slightly improves the explanatory power (model 3). The best model in terms of explained variance ($R^2 = 0.31$) among the linear ones is the model 4, which includes all factors and fixed effects. This model also have the lowest Akaike information criterion among the comparable linear models, meaning that the fit improvement is not due to overfitting. Model 5, which is a Poisson model with all variables is much a better fit regarding pseudo-$R^2$ and provides much more significant estimates (all fixed effects significant and much smaller standard deviations).

Across all models, consistent qualitative stylized facts regarding link creation between cities are revealed: (i) distance has a negative influence; (ii) economic sizes have a positive effect, with the size of the origin city being more important (consistent in the ownership links context, as strongly driven by decision at the headquarter/owner level); (iii) industrial proximity has a positive influence, i.e. similarity is more important than complementarity for this dimension; (iv) a large part of country fixed-effects are statistically significant; and (v) the order of magnitude of the influence of each factor is roughly the same, which means that no process totally dominates the others.
This confirms the relevance of including all these different dimensions in the generative network model, since they have a comparable explanatory role in statistical models.

@@

%%%%%%%%%%%%%
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{figures/Fig3.png}
    \end{center}
    \vspace{2cm}
    \caption{Example of simulated networks at $t=1500$, for a synthetic system of cities (Top row) and the European urban network (Bottom row), with a high (resp. low) gravity decay parameter for the left column (resp. right).\label{fig:fig3}}
\end{figure}
%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
	    \includegraphics[width=\linewidth]{figures/Fig4.png}
	\end{center}
	\vspace{2cm}
    \caption{(Top Left) Internationalisation index decreases exponentially with gravity decay; (Top Right) Correlation between city weighted degree and size. Both plots show a transition from a local to a global regime. (Bottom Left) Internationalisation varies linearly with sector proximity $\gamma_S$; (Bottom Right) Correlation between degree and size exhibits a maximum, witnessing an intermediate regime where size is the most important. \label{fig:fig4}}
\end{figure}
%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}
    \includegraphics[width=\linewidth]{figures/Fig5.png}
    \vspace{2cm}
    \caption{The transition as a function of interaction range depends on the influence of origin size $\gamma_O$; sector proximity $\gamma_S$ plays a role only for a large influence of the origin.\label{fig:fig5}}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}
    \includegraphics[width=\textwidth]{figures/Fig6.png}
    \vspace{2cm}
	\caption{Maximal integration in terms of community size is achieved at an intermediate value of d$_o$: emergence of a regional regime; maximal size depends on the role of sectors $\gamma_S$ in a decreasing way when the origin size is deactivated, and in an increasing way when $\gamma_o=1$; this regime disappears when the influence of the origin size becomes too large.\label{fig:fig6}}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
    \end{center}
    \vspace{2cm}
	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives. $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models). Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
\end{figure}
%%%%%%%%%%%%%%
